Association between trauma severity models and opportunities for improvement: A retrospective cohort study

### 2022-09-09 studyprogressreport.Rmd
method; development and time plan; lots of small fixes

TSS are already used in some clinics to identify cases for peer review [@whoguidelines]. However, the extent to which TSS are associated with or can predict opportunities for improvement (OFI) as found in M&M conferences is poorly understood. If it is found that an existing TSS model is associated with OFI, it could allow for another avenue of automatically identifying potentially mismanaged cases in a standardised fashion.

## Aims

This study aims to evaluate how trauma severity models are associated with and to what extent they can predict opportunities for improvement.

@@

<!-- Present key elements of study design -->
This is a single-centre retrospective cohort study, using data from two registries at Karolinska University Hospital (KUH): the trauma registry database and the trauma care quality registry. The trauma care quality registry is a subset of the trauma registry and contains data on trauma patients who were selected for morbidity and mortality (M&M) reviews. These two databases will be linked and then analysis will be performed to estimate the association between the common trauma severity models (TSM) TRISS, NORMIT, and RISCII, and opportunities for improvement (OFI) using logistic regression. The predictive performance of these models will be evaluated using measures of discrimination and calibration.

@@

The trauma registry at KUH contains data on 21,000 patients collected between 2012 and 2021. This registry is also submitted to SweTrau, Sweden's national trauma registry database, and therefore meets SweTrau guidelines. The registered data includes prehospital vital signs and management, vital signs and management on arrival at hospital, information about the type of injury, and the primary method of treatment. It also includes tracks follow-up. Follow-up variables include, but are not limited to, survival at 30 days, days in ventilator, GCS at discharge, highest level of care, and transfers to other units.

The trauma care quality register is an internal hospital register and registers the outcome of M&M conferences. M&M conferences are multidisciplinary and their purpose is to review selected patient cases and determine whether an opportunity for improvement existed in each individual case. Examples of opportunities for improvement could be insufficient resources or errors in patient management.

@@

<!-- Eligiblity criteria for the trauma registry were taken from the SWETRAU manual, check if it is exactly the same at? -->

@@

<!-- Update these numbers later when I have actual numbers -->

@@

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->

<!-- Really not sure
-->

@@

R, a programming language for statistical computing, will be used for compilation and analysis of data. Variables will be handled according to the SweTrau manual.

@@

All data required for this project were collected and available prior to the project beginning. Therefore, the data only need to be analysed.

>>>>

TSS are already used in some clinics to identify cases for peer review [@whoguidelines]. However, the extent to which TSS are associated with or can predict opportunities for improvement (OFI) as found in M&M conferences is poorly understood. If it is found that an existing TSS model is associated with OFI, it could allow for another avenue of automatically identifying potentially mismanaged cases in a standardised fashion. This study aims to evaluate how trauma severity models are associated with and to what extent they can predict opportunities for improvement.

@@

<!-- Present key elements of study design -->
This is a single-centre retrospective cohort study, using data from two registries at Karolinska University Hospital (KUH): the trauma registry database and the trauma care quality registry. The trauma care quality registry is a subset of the trauma registry and contains data on trauma patients who were selected for M&M reviews. These two databases will be linked and then analysis will be performed to estimate the association between the common trauma severity models (TSM) TRISS, NORMIT, and RISCII, and opportunities for improvement (OFI) using logistic regression. The predictive performance of these models will be evaluated using measures of discrimination and calibration.

@@

The trauma registry at KUH contains data on 21,000 patients collected between 2012 and 2021. This registry is also submitted to SweTrau, Sweden's national trauma registry database, and therefore meets SweTrau guidelines. The registered data includes prehospital vital signs and management, vital signs and management on arrival at hospital, information about the type of injury, and the primary method of treatment. Follow-up variables include, but are not limited to, survival at 30 days, days in ventilator, GCS at discharge, highest level of care, and transfers to other units.

The trauma care quality register is an internal hospital register and registers the outcome of M&M conferences. Each case reviewed is registered in the trauma care quality register with data about the outcome of the review.

@@

<!-- Eligiblity criteria for the trauma registry were taken from the SWETRAU manual, check if it is exactly the same at NKS? -->

@@

<!-- Update these numbers later when I have real numbers -->

@@

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
The exposure variables are the values given by the studied trauma models. These variables are continuous. The outcome variable, OFI, is a binary variable.

<!-- Really not sure exactly what to put here
-->

@@

R, a programming language for statistical computing, will be used for compilation and analysis of data. Conversion and handling of variables will be carried out according to the SweTrau manual. Results will be presented with a confidence interval of 95%. A p-value of <0.05 will be considered significant.

<!-- Complete case study? -->
<!--

@@

<!--
B) En skriftlig redogörelse på ca en A4-sida av projektets utveckling, plus tidsplan för
resterande del av kursen (använd tidsplanmallen i Canvas). Ange så exakt som möjligt:
• Vilken typ av data, kategorier, variabler etc du avser att samla in
• Hur data planeras presenteras (t ex tabeller och hur de kan struktureras, diagramtyper,
vävnadssnitt, intervjucitat etc)
• Hur data planeras analyseras; vilka statistiska metoder för jämförelser mellan en eller
flera grupper, andra matematiska metoder, kvalitativ analys etc
-->

This is a registry-based project and all data required were collected and available prior to the project beginning. The data consists of information about trauma patients at KUH and the outcomes of morbidity & mortality conferences on selected patient cases. The next step after submitting this progress report is to begin writing the code that will import and clean data. Once this is completed, the code for data analysis will be written. The analysis will consist of using logistic regression to see if trauma severity models can predict opportunities for improvement. Initially the code will be written using scrambled data in order to minimise researcher bias. Once the code for analysis is complete, it will be run on the real-life data to obtain the results for the study. I have written a detailed time plan, including dates for deadlines, that is available for my supervisors. We will be in contact on a regular basis to discuss progress. All of my code and text will be uploaded on a regular basis, which will allow for my supervisors to see my progress.
